---
title: "AdventureWorks Database Analysis"
author: "Quan Phuc Minh Nguyen"
date: "`r Sys.Date()`"
output:
  word_document: default
  pdf_document: default
  html_document: default
fontsize: 12pt
---

```{r SQL connection, include=FALSE}
library(RODBC)
con <- odbcConnect("mikesql",rows_at_time = 1)
```
## Introduction

The point of this analysis is to utilize the online and reseller sales data to determine which products or product categories keep customers coming back to AdventureWorks. The company also wants to identify which product generates the highest revenue per one product quantity.

Most data manipulation and transformation process will be done with SQL, while data visualization will be created with R and Tableau.

## Prepare

The prepare phase helps ensure data credibility and data integrity.

### Database location

The AdventureWorks database can be found [here](https://docs.microsoft.com/en-us/sql/samples/adventureworks-install-configure?view=sql-server-ver16&tabs=ssms).
 
### Duplicate and null values
Checking if the raw data is free of duplicate values allows the data to be unbiased and credible. The following SQL query checks for duplicate values in the reseller sales table - if a sales order number contains more than one product key, it is considered a duplicate.
```{r Duplicate, echo=TRUE}
Duplicate <-"
SELECT 
	SalesOrderNumber,
	ProductKey,
	COUNT(*)
FROM 
	AdventureWorksDW2019.dbo.FactResellerSales
GROUP BY
	SalesOrderNumber,
	ProductKey
HAVING
	COUNT(*)>1
;
"
```
```{r Duplicate Table, echo=FALSE}
sqlQuery(con,Duplicate,stringsAsFactors=FALSE)
```
We use a similar SQL query to check for duplicate values in the online sales and customer tables. Fortunately, based on the result, there are no duplicate orders and no duplicate customer information. 

### Non-subcategorized products
Some products do not have a subcategory key, so it is important to take into consideration the orders containing such products. From the table product, the first 209 products labeling 1-209 are not subcategorized. Thus, the following query returns orders with non-subcategorized products.
```{r Unsub, echo=TRUE}
Unsub <-"
SELECT 
	COALESCE(Res.ProductKey,Onl.ProductKey) AS ProductKey,
	COALESCE(Res.SalesOrderNumber,Onl.SalesOrderNumber) AS OrderNumber
FROM 
	AdventureWorksDW2019.dbo.FactResellerSales AS Res
	INNER JOIN
	AdventureWorksDW2019.dbo.FactInternetSales AS Onl 
	ON Res.ProductKey=Onl.ProductKey
WHERE
	COALESCE(Res.ProductKey,Onl.ProductKey)<210
;
"
```
```{r Unsub table, echo=FALSE}
sqlQuery(con,Unsub,stringsAsFactors=FALSE)
```
Results show that there are no such orders. Thus, we can move on to processing the data.

## Process
The process phase allows us to transform the data into meaningful patterns for analysis. 

### Best-selling products
The following SQL query returns the top 5 best-selling products of all time. This query also includes subqueries to find the top 5 best-selling products in each division.
```{r Best-selling product (BSP), echo=TRUE}
BestSellingProduct <-"
SELECT TOP 5
	EnglishProductName AS Product,
	TotalProductsPurchased AS QuantityPurchased
FROM
	(
		SELECT								
			COALESCE(Onl.ProductKey,Res.ProductKey) AS ProductKey,
			CASE
				WHEN Onl.QuantityPurchased IS NULL 
					THEN Res.QuantityPurchased
				WHEN Res.QuantityPurchased IS NULL 
					THEN Onl.QuantityPurchased
				ELSE Onl.QuantityPurchased+Res.QuantityPurchased 
			END AS TotalProductsPurchased
		FROM 
			(
				SELECT
					ProductKey,
					COUNT(ProductKey) AS QuantityPurchased
				FROM 
					AdventureWorksDW2019.dbo.FactInternetSales
				GROUP BY
					ProductKey
			) AS Onl -- Best-selling products online
			LEFT JOIN
			(
				SELECT
					ProductKey,
					COUNT(ProductKey) AS QuantityPurchased
				FROM 
					AdventureWorksDW2019.dbo.FactResellerSales
				GROUP BY
					ProductKey
			) AS Res -- Best-selling products from resellers
			ON Onl.ProductKey=Res.ProductKey
	) AS ProductSales
	LEFT JOIN AdventureWorksDW2019.dbo.DimProduct
	ON ProductSales.ProductKey=AdventureWorksDW2019.dbo.DimProduct.ProductKey -- Return the product name
ORDER BY
	TotalProductsPurchased DESC
;
"
```
```{r BSP table, echo=FALSE}
sqlQuery(con,BestSellingProduct,stringsAsFactors=FALSE)
```
The query first generates a table with product keys and the corresponding quantity. Then, it returns the products' names through a JOIN clause with the product description table.

### Best-selling categories
Each product has a category, and this SQL query reports the number of products purchased under each type.
```{r BSC, echo=TRUE}
BestSellingCategories <-"
SELECT
	COALESCE(Res.Category, Onl.Category) AS Category,
	COALESCE(Res.CategoriesPurchased,0) AS ResellerQuantity,
	COALESCE(Onl.CategoriesPurchased,0) AS OnlineQuantity
FROM
	(
		SELECT 
			EnglishProductCategoryName AS Category,
			COUNT(EnglishProductCategoryName) AS CategoriesPurchased
		FROM 
			(
				SELECT
					ProductKey,
					COUNT(ProductKey) AS QuantityPurchased
				FROM
					AdventureWorksDW2019.dbo.FactResellerSales
				GROUP BY
					ProductKey
			) AS TotalNoProducts
				LEFT JOIN
					AdventureWorksDW2019.dbo.DimProduct
					ON TotalNoProducts.ProductKey=AdventureWorksDW2019.dbo.DimProduct.ProductKey
				LEFT JOIN
					AdventureWorksDW2019.dbo.DimProductSubcategory
					ON AdventureWorksDW2019.dbo.DimProduct.ProductSubcategoryKey
					=AdventureWorksDW2019.dbo.DimProductSubcategory.ProductSubcategoryKey
				LEFT JOIN
					AdventureWorksDW2019.dbo.DimProductCategory
					ON AdventureWorksDW2019.dbo.DimProductSubcategory.ProductCategoryKey
					=AdventureWorksDW2019.dbo.DimProductCategory.ProductCategoryKey
		GROUP BY
			EnglishProductCategoryName
	) AS Res
	FULL OUTER JOIN
	(
		SELECT 
			EnglishProductCategoryName AS Category,
			COUNT(EnglishProductCategoryName) AS CategoriesPurchased
		FROM 
			(
				SELECT
					ProductKey,
					COUNT(ProductKey) AS QuantityPurchased
				FROM 
					AdventureWorksDW2019.dbo.FactInternetSales
				GROUP BY
					ProductKey
			) AS TotalNoProducts
			LEFT JOIN
				AdventureWorksDW2019.dbo.DimProduct
				ON TotalNoProducts.ProductKey=AdventureWorksDW2019.dbo.DimProduct.ProductKey
			LEFT JOIN
				AdventureWorksDW2019.dbo.DimProductSubcategory
				ON AdventureWorksDW2019.dbo.DimProduct.ProductSubcategoryKey
				=AdventureWorksDW2019.dbo.DimProductSubcategory.ProductSubcategoryKey
			LEFT JOIN
				AdventureWorksDW2019.dbo.DimProductCategory
				ON AdventureWorksDW2019.dbo.DimProductSubcategory.ProductCategoryKey
				=AdventureWorksDW2019.dbo.DimProductCategory.ProductCategoryKey
		GROUP BY
			EnglishProductCategoryName
	) AS Onl ON Res.Category=Onl.Category
;
"
```
```{r BSC table, echo=FALSE}
sqlQuery(con,BestSellingCategories,stringsAsFactors=FALSE)
```

### Best-selling subcategories
We want to further examine each category and find out the exact quantity under each subcategory. Suppose we want to acquire a table with quantity count for each subcategory under the category Bikes that are purchased online. Similar to the previous query, the following one generates the quantity from the sales table and connects to the product tables to return the names for Bikes' subcategories.
```{r Online Bikes, echo=TRUE}
OnlineBikes <-"
SELECT 
	EnglishProductCategoryName AS Category,
	EnglishProductSubcategoryName AS Subcategory,
	COUNT(EnglishProductSubcategoryName) AS QuantitiesPurchasedOnline
FROM 
	(
		SELECT
			ProductKey,
			COUNT(ProductKey) AS QuantityPurchased
		FROM 
			AdventureWorksDW2019.dbo.FactInternetSales
		GROUP BY
			ProductKey
	) AS Onl
	LEFT JOIN
		AdventureWorksDW2019.dbo.DimProduct
		ON Onl.ProductKey=AdventureWorksDW2019.dbo.DimProduct.ProductKey
	LEFT JOIN
		AdventureWorksDW2019.dbo.DimProductSubcategory
		ON AdventureWorksDW2019.dbo.DimProduct.ProductSubcategoryKey
		=AdventureWorksDW2019.dbo.DimProductSubcategory.ProductSubcategoryKey
	LEFT JOIN
		AdventureWorksDW2019.dbo.DimProductCategory
		ON AdventureWorksDW2019.dbo.DimProductSubcategory.ProductCategoryKey
		=AdventureWorksDW2019.dbo.DimProductCategory.ProductCategoryKey
WHERE 
	EnglishProductCategoryName='Bikes'
GROUP BY
	EnglishProductCategoryName, 
	EnglishProductSubcategoryName
;
"
```
```{r OB Table, echo=FALSE}
sqlQuery(con,OnlineBikes,stringsAsFactors=FALSE)
```
### Total Sales
We would like to analyze total sales to identify patterns and aspects of sales AdventureWorks wants to address. The following SQL query retrieves the sales data in the U.S. territory from 2011 to 2013.
```{r TS, echo=TRUE}
TotalSales <-"
WITH
	Res AS (
		SELECT 
			CASE 
				WHEN CAST(SUBSTRING(CAST(OrderDateKey AS CHAR),5,2) AS INT) BETWEEN 1 AND 3 
					THEN LEFT(OrderDateKey, 4) + ' Q1'
				WHEN CAST(SUBSTRING(CAST(OrderDateKey AS CHAR),5,2) AS INT) BETWEEN 4 AND 6  
					THEN LEFT(OrderDateKey, 4) + ' Q2'
				WHEN CAST(SUBSTRING(CAST(OrderDateKey AS CHAR),5,2) AS INT) BETWEEN 7 AND 9 
					THEN LEFT(OrderDateKey, 4) + ' Q3'
				ELSE LEFT(OrderDateKey, 4) + ' Q4'
			END AS SalesQuarter,
			ROUND(SUM(SalesAmount),0) AS Sales
		FROM AdventureWorksDW2019.dbo.FactResellerSales
		WHERE SalesTerritoryKey BETWEEN 1 AND 5
		GROUP BY OrderDateKey
	),
	Onl AS (	
		SELECT 
			CASE 
				WHEN CAST(SUBSTRING(CAST(OrderDateKey AS CHAR),5,2) AS INT) BETWEEN 1 AND 3 
					THEN LEFT(OrderDateKey, 4) + ' Q1'
				WHEN CAST(SUBSTRING(CAST(OrderDateKey AS CHAR),5,2) AS INT) BETWEEN 4 AND 6  
					THEN LEFT(OrderDateKey, 4) + ' Q2'
				WHEN CAST(SUBSTRING(CAST(OrderDateKey AS CHAR),5,2) AS INT) BETWEEN 7 AND 9 
					THEN LEFT(OrderDateKey, 4) + ' Q3'
				ELSE LEFT(OrderDateKey, 4) + ' Q4'
			END AS SalesQuarter,
			ROUND(SUM(SalesAmount),0) AS Sales
		FROM AdventureWorksDW2019.dbo.FactInternetSales
		WHERE SalesTerritoryKey BETWEEN 1 AND 5
		GROUP BY OrderDateKey
	)
SELECT 
	COALESCE(Res.SalesQuarter,Onl.SalesQuarter) AS SalesPeriod,
	COALESCE(SUM(Res.Sales), 0) AS ResellerSales,
	COALESCE(SUM(Onl.Sales), 0) AS OnlineSales,
	COALESCE(SUM(Onl.Sales), 0)+COALESCE(SUM(Res.Sales), 0) AS TotalSales
FROM 
	Res 
	FULL OUTER JOIN
	Onl ON Res.SalesQuarter=Onl.SalesQuarter
WHERE LEFT(COALESCE(Res.SalesQuarter,Onl.SalesQuarter),4) BETWEEN 2011 AND 2013
GROUP BY COALESCE(Res.SalesQuarter,Onl.SalesQuarter)
ORDER BY SalesPeriod;
"
```
```{r TS table, echo=FALSE}
sqlQuery(con,TotalSales,stringsAsFactors=FALSE)
```
## Analysis
### Visualization
Data visualization is helpful for identifying patterns and trends. From the TotalSales table, online sales is relatively low compared to the total sales, so we illustrate the data in two bar graphs. 

#### Reseller vs. Total & Online
Since the sales data is a discrete variable, it can be ideally represented by a bar chart. The table generated by SQL is a wide-formatted data frame, so we need to reshape it into a long-formatted one for the sake of using R. Using the tidyr package: 
```{r Importing SQL tables, include=FALSE}
library(readr)
library(dplyr)
TotalSales<- read_csv("E:\\AdventureWorks Project\\totalsales.csv")
Category <- read_csv("E:\\AdventureWorks Project\\Category.csv")
```
```{r Reshaping format, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyr)
TotalSalesLong <- TotalSales %>% pivot_longer(cols=c('ResellerSales','OnlineSales','Sales'),
                                              names_to='Type',
                                              values_to='SalesAmount')
```
Filtering out the online sales data:
```{r}
RvT <- TotalSalesLong %>% filter(Type=='ResellerSales'|Type=='Sales')
```
Now, we can plot our data using ggplot2 package:
```{r RvT, echo=TRUE, message=FALSE, warning=FALSE}
library(ggplot2)
library(scales)
library(tidyverse)

ggplot(data=RvT)+
  geom_bar(mapping=aes(x=SalesPeriod,y=SalesAmount,fill=Type,width=.5),stat='identity',position='dodge')+
  scale_y_continuous(labels = unit_format(unit = "M", scale = 1e-6))+ # Rescaling metrics
  scale_fill_manual(values = c("#F95700","#00539C"))+ # Contrasting color pallete
  labs(title='Quarterly Reseller vs. Quarterly Total Sales From 2011 To 2013')+
  theme(axis.text.x=element_text(angle=45)) # Make the x-axis readable
```
```{r Online Sales, echo=TRUE}
ggplot(data=TotalSales)+
  geom_bar(aes(x=SalesPeriod,y=OnlineSales),stat='identity',fill="#00539C")+
  scale_y_continuous(labels = unit_format(unit = "M", scale = 1e-6))+
  labs(title='Quarterly Online Sales From 2011 To 2013')+
  theme(axis.text.x=element_text(angle=45))
```

#### Distribution of Categories
The distribution is best illustrated by a pie chart.
```{r Categories Dist., echo=TRUE, message=FALSE, warning=FALSE}
ggplot(data=Category,aes(x=" ",y=Quantity, fill=Category))+
  geom_bar(stat="identity",color="black")+
  coord_polar("y",start=0)+
  theme_void()+
  geom_text(aes(label = paste0(round(Quantity/sum(Quantity)*100), "%")), 
            position = position_stack(vjust = 0.5))+
  labs(title='Category Distribution')
```

### Sales insights
In the grouped bar chart, we can notice varying demand patterns within a year, but the average annual sales amount increased every year. Hence, meticulous planning and a flexible manufacturing system would be an excellent solution for AdventureWorks. AdventureWorks should allocate more resources during leaner times of the year for the pre-manufacturing work like the production of components and packages. On the other hand, during busier times, it should focus on manufacturing and distributing products. Since the database only provides two years of sales data, more pattern analysis might be required to foster more effective allocation decisions.

From the online sales chart, we can also see a fluctuating pattern within two years. However, the average sales amount significantly increased in 2013, potentially due to technology and the online platform market improvement. Hence, although the online sales amount was remarkably lower, AdventureWorks should devote more resources to enhancing its online platform, like launching advertising campaigns or building its market app. With more technology becoming readily available, the online platform should be the primary source of sales in no time.

Product category analysis and prospective customer insights can help increase sales and satisfaction and allocate resources to the proper manufacturing sector. The pie plot indicates that components and bikes account for most of the sales. AdventureWorks could utilize this information by planning for product improvements and better promotional strategies.

